### 2018-06-01 source/Ch7-loresEGS/lowres-EGS.tex
added a lot

\pg
In this section, we describe the work done to make a wide-field image of the entire LOFAR primary beam, centred on EGS. Such a large-field, low-frequency image of this field has never yet been done in this frequency band. Due to technical limitations, the maximum attainable image size (in terms of number of pixels) will limit resolution to about $5''$. As such, the international stations are not used for this part of the project.

\pg
A more complete sky model can be created from this image, allowing for one last round of self-calibration on the 3C295 high-resolution model. Bright out-of-field sources can be subtracted from images of the Extended Groth Strip so that their sidelobes do not contaminate the final images. To achieve this, we require direction-dependent calibration. Because we do not use the international LOFAR stations, we do not use our high-resolution model of 3C295. This ensures that we do not ``hard-set" errors in our high-resolution model caused by other sources in the field.

\pg
Our methodology is therefore to calibrate the full LOFAR bandwidth using the LOFAR Surveys KSP pipeline (talk about DD calibration, facetting, etc). We then perform source analysis on the most interesting objects in the field, using ancillary data from NVSS \citepads[NRAO VLA Sky Survey, ]{1998AJ....115.1693C}, SDSS \citepads[Sloan Digital Sky Survey, ]{2000AJ....120.1579Y} and WISE \citepads{2010AJ....140.1868W} surveys and cross-referenced using SIMBAD \citepads{2000A&AS..143....9W}. Overlays for those sources will be shown in \cref{sec.lowresEGS.overlays}.

@@

\subsection{Calibration Strategy}

\pg
An initial sky model is generated using the VLSSr \citepads[VLA Low-frequency Sky Survey - redux, ]{2012RaSc...47.0K04L}, WENSS \citepads[WEsterbork Northern Sky Survey]{1997A&AS..124..259R} and the NVSS \citepads[NRAO VLA Sky Survey, ]{1998AJ....115.1693C}. This model serves as the basis for facet-based self-calibration, performed using the killMS-DDF surveys pipeline \citepads{2017A&A...598A.104S}.

\begin{figure}[h!]
\centering
\includegraphics[width=0.95\textwidth]{images/{EGS.widefield.regions}.png}
\caption{\label{fig.EGS.widefield.regions} Full direction-dependent calibrated image of the LOFAR primary beam. Image made without international stations.}
\end{figure}

\pg
The direction-dependent calibration was done using 121 facets, spread as shown in \cref{fig.EGS.widefield.regions}. Some facets converge faster or better than others. Calibration is done solving for full, complex Jones matrices, rather than a physics-based or diagonal-and-rotation matrix [CHECK THAT THIS IS TRUE!].
The final image was imaged with a very small $uv$-cut, keeping only baselines longer than 100m. This is because shorter baselines were not used during calibration, and so discarded for imaging. The image resolution is $1.5''$, and the image was created using both quality-based weighting schemes and Briggs weighting, with a Briggs factor of $-0.5$. This means that the imaging was skewed slightly in favour of uniform weighting (weighting visibilities by local $uv$-density) rather than natural weighting (same weight for all visibilities). Imaging and calibration were done modelling the LOFAR beam, so the image gives integrated flux values rather than apparent flux values. Deconvolution was performed using SSD-clean with an external mask.

\pg
\textcolor{red}{show self-cal loop (before-after) in images \& gains; show improvement from direction-dependent cal. on entire field \& individual sources}

@@

\subsection{Selected Sources}

@@

\subsection{Images \& Cross-matching}

\pg
We will now show images, overlays and cross-matching analysis \& results for the 12 chosen sources in the primary beam.

@@

\begin{figure}[h!]
\centering
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs1.LOFARegs}.png}}
\caption{\label{fig.egs1.lofarim} Cutout of our LOFAR wide-field image centred on EGS-1.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs1.mosaic.fits.lofar}.png}}
\caption{\label{fig.egs1.sdss.lofoverlay} SDSS image of EGS-1, with \cref{fig.egs1.lofarim} as an overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs1.mosaic.fits.nvss}.png}}
\caption{\label{fig.egs1.sdss.nvssoverlay} Same image as \cref{fig.egs1.sdss.lofoverlay}, with NVSS overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs1.mosaic.fits.Wise3.4}.png}}
\caption{\label{fig.egs1.sdss.wiseoverlay} Same image as \cref{fig.egs1.sdss.lofoverlay}, with WISE overlay.}
\end{subfigure}
\caption{\label{fig.egs1} Images \& Overlays for EGS-1.}
\end{figure}

\pg
EGS-1 was selected from the wide-field LOFAR image because it showed interesting diffuse structure. \cref{fig.egs1} shows four images: a cutout of the LOFAR image around the source, an SDSS cutout of the same area with the LOFAR image superimposed as an overlay, the same SDSS cutout with an NVSS image overlayed, and finally the same SDSS image again with a WISE image overlay. The first image aims to show why the source was selected. The second shows optical counterparts to LOFAR emission. The third serves to check whether we pick up similar structure and emission in NVSS images as the LOFAR wide-field image. Note that the NVSS images are at a much lower resolution, and are therefore not expected to give any information on LOFAR source structure, but can be a useful way to ensure that we are not choosing artefacts or spurious emission as sources to investigate. Finally, the WISE overlay attempts to see whether infrared emission can be associated to these sources. Note that the sensitivity of SDSS images are not homogeneous: this is because not all patches of the sky have received the same coverage.

\pg
EGS-1 does not seem to match up to any optical source, nor does it match an infrared source. EGS-1 therefore seems to be best identified as a low-frequency counterpart to NVSS J143706+533225. Its shape seems to include two jets, one much brighter than the other, but both visible in \cref{fig.egs1.lofarim}. This is congruent with a radio galaxy with one Doppler-boosted jet. The absence of an optical counterpart could be due to dust extinction or to insufficient sensitivity.

%\clearpage
\subsubsection{EGS-2}

\begin{figure}[h!]
\centering
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs2.LOFARegs}.png}}
\caption{\label{fig.egs2.lofarim} Cutout of our LOFAR wide-field image centred on EGS-2.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs2.mosaic.fits.lofar}.png}}
\caption{\label{fig.egs2.sdss.lofoverlay} SDSS image of EGS-2, with \cref{fig.egs2.lofarim} as an overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs2.mosaic.fits.nvss}.png}}
\caption{\label{fig.egs2.sdss.nvssoverlay} Same image as \cref{fig.egs2.sdss.lofoverlay}, with NVSS overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs2.mosaic.fits.Wise3.4}.png}}
\caption{\label{fig.egs2.sdss.wiseoverlay} Same image as \cref{fig.egs2.sdss.lofoverlay}, with WISE overlay.}
\end{subfigure}
\caption{\label{fig.egs2} Images \& Overlays for EGS-2.}
\end{figure}

\pg
EGS-2 was selected from the wide-field LOFAR image because of the interestingly turbulent morphology of its northern jet, and extended diffuse emission. \cref{fig.egs2} shows four images: a cutout of the LOFAR image around the source, an SDSS cutout of the same area with the LOFAR image superimposed as an overlay, the same SDSS cutout with an NVSS image overlayed, and finally the same SDSS image again with a WISE image overlay. The first image aims to show why the source was selected. The second shows optical counterparts to LOFAR emission. The third serves to check whether we pick up similar structure and emission in NVSS images as the LOFAR wide-field image. Note that the NVSS images are at a much lower resolution, and are therefore not expected to give any information on LOFAR source structure, but can be a useful way to ensure that we are not choosing artefacts or spurious emission as sources to investigate. Finally, the WISE overlay attempts to see whether infrared emission can be associated to these sources. Note that the sensitivity of SDSS images are not homogeneous: this is because not all patches of the sky have received the same coverage.

\pg
A bright optical source lies very close to the mid-point of EGS-2. This source - 2MASX J14352846+5507519\footnote{\url{http://simbad.u-strasbg.fr/simbad/sim-id?Ident=\%40499314&Name=2MASX\%20J14352846\%2b5507519&submit=submit}} - is the brightest galaxy in a cluster. It lies at a redshift of 0.13985 \citepads{2013yCat.5139....0A}. Furthermore, there is an NVSS source which seems to nicely overlap the EGS-2: NVSS J143527+550756 \citepads{1998AJ....115.1693C}. These two sources do not seem to be associated in the literature. [reasons why they aren't?]

\pg
The northern edge of EGS-2's upper jet has interesting morphology: if EGS-2 is indeed associated to the brightest galaxy in the cluster, then this morphology is likely due to differing densities of the intercluster and intracluster media, which the jets will penetrate and interact with in different ways. This morphology is lost on the overlay, and so may not be statistically significant enough to draw conclusions from, however. As ever, because this source lies far from the observation pointing, the sensitivity and calibration are not optimal for analysis: further pointings on EGS-2 might be required to achieve the signal-to-noise required to extract useful information, especially using international LOFAR.

%\clearpage
\subsubsection{EGS-3}

\pg
EGS-3 is one of the most impressive sources in the primary beam.It has very complex diffuse structure. It almost certainly consists of two jets, though the interactions of these jets with the neighbouring medium must be complex indeed. From \cref{fig.egs3.sdss.nvssoverlay}, one can see that the emission can be successfully matched to no less than two individual NVSS sources. In other words, if those NVSS sources are linked to the LOFAR source - and given their spatial distribution, they almost certainly are - then they are two components of a single, deeper structure. The infrared image tells us very little.

\begin{figure}[h!]
\centering
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs3.LOFARegs}.png}}
\caption{\label{fig.egs3.lofarim} Cutout of our LOFAR wide-field image centred on EGS-3.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs3.mosaic.fits.lofar}.png}}
\caption{\label{fig.egs3.sdss.lofoverlay} SDSS image of EGS-3, with \cref{fig.egs3.lofarim} as an overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs3.mosaic.fits.nvss}.png}}
\caption{\label{fig.egs3.sdss.nvssoverlay} Same image as \cref{fig.egs3.sdss.lofoverlay}, with NVSS overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs3.mosaic.fits.Wise3.4}.png}}
\caption{\label{fig.egs3.sdss.wiseoverlay} Same image as \cref{fig.egs3.sdss.lofoverlay}, with WISE overlay.}
\end{subfigure}
\caption{\label{fig.egs3} Images \& Overlays for EGS-3.}
\end{figure}

@@

\begin{figure}[h!]
\centering
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs4.LOFARegs}.png}}
\caption{\label{fig.egs4.lofarim} Cutout of our LOFAR wide-field image centred on EGS-4.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs4.mosaic.fits.lofar}.png}}
\caption{\label{fig.egs4.sdss.lofoverlay} SDSS image of EGS-4, with \cref{fig.egs4.lofarim} as an overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs4.mosaic.fits.nvss}.png}}
\caption{\label{fig.egs4.sdss.nvssoverlay} Same image as \cref{fig.egs4.sdss.lofoverlay}, with NVSS overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs4.mosaic.fits.Wise3.4}.png}}
\caption{\label{fig.egs4.sdss.wiseoverlay} Same image as \cref{fig.egs4.sdss.lofoverlay}, with WISE overlay.}
\end{subfigure}
\caption{\label{fig.egs4} Images \& Overlays for EGS-4.}
\end{figure}

@@

\pg
The work done in this section had for primary goal to create a decent model of the sky seen by LOFAR's primary beam. To do this, a wide-field, low-resolution image of the fully primary beam was made: this image makes no use of the LOFAR international stations due to technical limitations. Using PyBDSM \citepads{2015ascl.soft02007M}, this model was extracted from an image made after a few rounds of direction-dependent self-calibration.

>>>>

\pg
In this section, we describe the work done to make a wide-field image of the entire LOFAR primary beam, centred on EGS. A large-field, image of this field has never yet been made at such low frequencies. Due to technical limitations, the maximum attainable image size (in terms of number of pixels) will limit resolution to about $5''$. As such, the international stations are not used for this part of the project.
A more complete sky model can be created from this image, and %, allowing for one last round of self-calibration on the 3C295 high-resolution model. 
bright out-of-field sources can be subtracted from high-resolution images of the Extended Groth Strip so that their sidelobe emission do not contaminate the final images. To achieve this, we require direction-dependent calibration. Because we do not use the international LOFAR stations, we do not use our high-resolution model of 3C295. %  This ensures that we do not ``hard-set" errors in our high-resolution model caused by other sources in the field.

\pg
We aim to achieve a sensitivity of a few hundreds of $\mu$Jy using visibilities recorded between $110-180$ MHz. This corresponds to a sensitivity of $\sim20$ $\mu$Jy at 1.4 GHz for typical synchrotron radio sources, which have spectral indices of $\alpha\sim -0.7$ (where $S_\nu \propto \nu^\alpha$). Reaching this sensitivity goal over the full LOFAR primary beam requires direction-dependent calibration.
Our methodology is therefore to calibrate the full LOFAR bandwidth using the LOFAR Surveys KSP pipeline, which performs a facet-based direction-dependent calibration. We then perform source analysis on the most interesting objects in the field, using ancillary data from NVSS \citepads[NRAO VLA Sky Survey, ]{1998AJ....115.1693C}, SDSS \citepads[Sloan Digital Sky Survey, ]{2000AJ....120.1579Y} and WISE \citepads{2010AJ....140.1868W} surveys and cross-referenced using SIMBAD \citepads{2000A&AS..143....9W}. Overlays for those sources will be shown in \cref{sec.lowresEGS.overlays}.

@@

%\subsection{Calibration Strategy}

\pg
Reducing LOFAR data is a complex business for a number of reasons. When seeking to make high-fidelity, wide-field images, the need to minimise decorrelation (cf. \cref{section.RIME.TimeDep.Decoherence}) means that data can only be averaged with care. As visibilities are averaged in time and frequency\footnote{Provided that simple averaging is performed; the use of baseline-dependent window functions can alleviate this effect to some extent}, the relative contribution of signal picked up on longer baselines from sources far from the observation centre is decreased. This translates to a blurring of sources as distance from phase centre increases.

\pg
What's more, antenna gains can vary over short durations or show peculiar spectral behaviour. Even if interferometric data were averaged massively using baseline-dependent window functions, information on this underlying gain behaviour would be lost - this means that averaging data to reduce its size (and associated reduction wall time) is a risky proposition.

\pg
As such, we seek to reduce an eight-hour LOFAR HBA observation, averaged down to 1 measurement per second and per 24 kHz\footnote{This is the bandwidth for each channel of our observation.}. This corresponds to a total of 7 terabytes of raw data: 2920 frequency measurements each second, for 28800 second, of 4 correlations, for each baseline. The observation mode used was HBA\textunderscore DUAL\textunderscore INNER, in which all Dutch stations operate with 24 of their 48 tiles, and these substations are then correlated separately with the rest of the array. This was chosen because this configuration does not negatively impact the array density of short baselines (allows for better sensitivity to, and thus recovery of, diffuse emission), nor does it introduce the calibration difficulties that non-uniform beams would cause.

\pg
The calibration strategy consists of a first round of direction-independent calibration followed by direction-dependent self-calibration. In effect, we use the following RIME:
\begin{equation}
\Vis_{pq} = \Gjones_p \bigg( \sum_s \Ejones_{sp} \Kjones_{sp} \Bmatrix_{s} \Kjones_{sq}^H \Ejones_{sq}^H \bigg) \Gjones_q^H
\end{equation}
which is the same as \cref{eq.multiple.source.RIME}. To account for the faceting explicitly, we can write:
\begin{equation}
\Vis_{pq} = \Gjones_p \sum_{\Omega}\bigg( \sum_{s\in\Omega} \Ejones_{sp} \Kjones_{sp} \Bmatrix_{s} \Kjones_{sq}^H \Ejones_{sq}^H \bigg) \Gjones_q^H
\end{equation}
where $\Omega$ indices denote different facets, and $s$ indices denote individual sources. Each source in a given facet is assumed to have similar direction-dependent gains, and their cumulative signal-to-noise allows for better estimations of this direction-dependent gain than would be possible if solving for each source individually. Thus:
\begin{equation}
\Vis_{pq} = \Gjones_p \sum_{\Omega}\bigg( \sum_{s\in\Omega} \Ejones_{\Omega p} \Kjones_{sp} \Bmatrix_{s} \Kjones_{sq}^H \Ejones_{\Omega q}^H \bigg) \Gjones_q^H
\end{equation}

\pg
We begin with direction-independent calibration to find an estimate of $\Gjones_p\forall p$. This first step is done because this part of the Jones chain is the same throughout the sky, and its inverse solution can therefore be applied directly to the visibilities as follows:
\begin{align}
\Vis_{pq}^\mathrm{corr} &= \Ghatjones_p^{-1} \Vis_{pq} \left( \Ghatjones_q^H \right)^{-1}\\
						&= \Ghatjones_p^{-1}\Gjones_p \sum_{\Omega}\bigg( \sum_{s\in\Omega} \Ejones_{\Omega p} \Kjones_{sp} \Bmatrix_{s} \Kjones_{\Omega q}^H \Ejones_{sq}^H \bigg) \Gjones_q^H \left( \Ghatjones_q^H \right)^{-1}
\end{align}
In so doing, we assume that $\left(\Ejones_{sp}\forall s,p\right) =1$ when solving for $\Ghatjones$. In general, this is of course invalid. For the calibrator source, however, we have $Ejones_{sp} =1 \forall p$ by construction. Because we can choose the calibrator source, the presence of a sufficiently bright source (in our case, an unresolved 3C295) will give extremely good estimations of $\Gjones$. We then have $\Ghatjones_p^{-1}\Gjones_p\sim 1$, and
\begin{align}
\Vis_{pq}^\mathrm{corr} &\sim \sum_{\Omega}\left( \sum_{s\in\Omega} \Ejones_{\Omega p} \Kjones_{sp} \Bmatrix_{s} \Kjones_{sq}^H \Ejones_{\Omega q}^H \right)
\end{align}
and direction-dependent calibration then consists of solving for $\Ehatjones_{\Omega p} = \Ghatjones_p^{-1}\Gjones_p\Ejones_{\Omega p} \sim \Ejones_{\Omega p}$ for all facets.

\pg
Direction-independent calibration is done with Prefactor \citepads{2016ApJS..223....2V}, using an initial sky model generated using the VLSSr \citepads[VLA Low-frequency Sky Survey - redux, ]{2012RaSc...47.0K04L}, WENSS \citepads[WEsterbork Northern Sky Survey]{1997A&AS..124..259R} and the NVSS \citepads[NRAO VLA Sky Survey, ]{1998AJ....115.1693C}, with 3C295 dominating the model. This model also serves as the initial model for the direction-dependent facet-based self-calibration on the corrected visibilities, performed using the killMS-DDF pipeline \citepads{2017A&A...598A.104S}. The quality-based weighting scheme discussed in \cref{chapter.paper} was used to improve self-calibration convergence and the final image.

\pg
Direction-dependent calibration was required, as the field of view is large enough that differential ionospheric errors are introduced as a function of direction: the signal from different sources cross different sections of the ionosphere, and the actual gains for these sources are thus different from the gains for the calibrator. They must thus be solved for individually. Assuming that the scale of ionospheric change is of the order of a degree (which is what is observed empirically), then a facet-based approach will give a ``good enough" result for most of the field. The direction-dependent calibration consisted of a physics-based approach, solving for XX and YY correlations along with a rotation matrix simulating differential Faraday rotation. These gain solutions were then smoothed in time and frequency to provide a time-independent, frequency-dependent amplitude solution for each station. The full LOFAR bandwidth is large enough to allow for Clock/TEC separation, which consists of explicitly modeling the two direction-independent phenomena which dominate in the gain's spectral structure. The clock difference between stations\footnote{Each station has a relatively-accurate clock associated with it, save for the Superterp stations which all share a single clock. Like all clocks, they have finite accuracy, and desynchronisation errors thus creep in over time.} introduces a phase error proportional to $\nu$, while differences in Total Electron Content (TEC) along the line of sight of different stations introduce a phase error inversely proportional to $\nu$. With sufficient spectral coverage, these two effects can therefore be separated and modeled individually, thereby improving the phase solution conditioning.

\begin{figure}[h!]
\centering
\includegraphics[width=0.95\textwidth]{images/{EGS.widefield.regions}.png}
\caption{\label{fig.EGS.widefield.regions} Facet distribution for the full direction-dependent calibrated image of the LOFAR primary beam. Image made without international stations.}
\end{figure}

\pg
The direction-dependent calibration was done using 121 facets, spread as shown in \cref{fig.EGS.widefield.regions}. Some facets converge faster or better than others, and so artefacts due to direction-dependent effects may remain in certain areas of the image but not others. Calibration is done solving for full, complex Jones matrices, rather than a physics-based or diagonal-and-rotation matrix.
The final image was imaged with a $uv$-cut, keeping only baselines longer than 100m. This is because shorter baselines were not used during calibration, and so discarded for imaging. The image resolution is $1.5''$, and the image was created using both quality-based weighting schemes and Briggs weighting, with a Briggs factor of $-0.5$. This means that the imaging was skewed slightly in favour of uniform weighting (weighting visibilities by local $uv$-density) rather than natural weighting (same weight for all visibilities). Imaging and calibration were done using the LOFAR beam model used in AWImager \citepads{awimager}, so the image gives integrated flux values rather than apparent flux values. Deconvolution was performed on the full bandwidth\footnote{This improves the conditioning of the imaging inverse problem, as it means that bright structure present across the full band (e.g. hotspots) can be deconvolved with the smaller PSF peak of the hightest-frequency band.} using SSD-clean with an external mask.

\begin{figure}[h!]
\centering
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{image_dirin_SSD_m_c_di_m.int.restored.fits}.png}}
\caption{\label{image.egs.widefield.noddsc} LOFAR wide-field image cutout, calibrated with the initial sky model and no direction-dependent Jones terms.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{image_full_ampphase_di_m.NS.app.restored.fits}.png}}
\caption{\label{image.egs.widefield.doddsc} LOFAR wide-field image cutout, self-calibrated from \cref{image.egs.widefield.noddsc} with direction-dependent Jones terms.}
\end{subfigure}
\caption{\label{image.egs.widefield.ddsc} Wide-field image cutouts of the full primary beam, centred on the EGS, before and after selfcal.}
\end{figure}

\pg
Direction-dependent self-calibration dramatically improves the noise level in the final image, as shown in \cref{image.egs.widefield.ddsc}. Note that not only do artefacts centred on 3C295 (which dominate the noise level in \cref{image.egs.widefield.noddsc}) decrease dramatically, but many of the "diffuse" sources in the field also seem to disappear: this is because their associated artefacts also decrease dramatically, and only physical emission remains. % Because this is a very large and high-resolution image, this means that most of these sources end up occupying a single pixel of the png cutout, and so appear to disappear.

\begin{figure}[h!]
\centering
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{widefield.DI}.png}}
\caption{\label{image.egs.widefield.noddsc.zoom} LOFAR wide-field image cutout, calibrated wtih the initial sky model and no direction-dependent Jones terms.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{widefiled.DD}.png}}
\caption{\label{image.egs.widefield.doddsc.zoom} LOFAR wide-field image cutout, self-calibrated from \cref{image.egs.widefield.noddsc} with direction-dependent Jones terms.}
\end{subfigure}
\caption{\label{image.egs.widefield.ddsc.zoom} Cutouts of smaller regions of \cref{image.egs.widefield.ddsc}.}
\end{figure}

\pg
These sources nevertheless remain in the field, as shown in \cref{image.egs.widefield.ddsc.zoom}. Note that the noise level is not constant throughout the image: this is because the array is not uniformly sensitive to the entire sky, but has a specific beam response. Correcting for this beam response in the image-plane gives integrated flux rather than apparent flux. % gives rise to this effect.

@@

%\subsection{Selected Sources}

@@

\pg
\cref{table.egs.sources} also gives the likely associations for each chosen EGS source, giving redshift information when available. The analysis here is done by hand using NED (\href{https://ned.ipac.caltech.edu/}{NASA/IPAC Extragalactic Database (NED)}), Simbad (\href{http://simbad.u-strasbg.fr/simbad/}{
operated at CDS, Strasbourg}) and the associated catalogues and papers, often accessible through ADS (\href{http://adsabs.harvard.edu/}{SAO/NASA Astrophysics Data System (ADS)}). Such an artisanal approach is suitable for the analysis of a dozen sources, but quickly becomes intractable. % Ideally, one would rely on automated matching between sources extracted from the widefield image and complete catalogues at other wavelengths: this would allow the matching of hundreds of sources in the time it took to analyse 12.
Since source analysis is but a happy byproduct of our primary science aim, this was not done here, but would be a promising avenue of future work.

\pg
For each source, we show four images: a cutout of the LOFAR image around the source, an SDSS cutout of the same area with the LOFAR image superimposed as an overlay, the same SDSS cutout with an NVSS image overlayed, and finally the same SDSS image again with a WISE image overlay. The first image aims to show why the source was selected. The second shows optical counterparts to LOFAR emission. The third serves to check whether we pick up similar structure and emission in NVSS images as the LOFAR wide-field image. Note that the NVSS images are at a much lower resolution, but can be a useful way to ensure that we are not choosing artefacts or spurious emission as sources to investigate. Finally, the WISE overlay attempts to see whether infrared emission can be associated to these sources. Note that the sensitivity of SDSS images are not homogeneous, as not all patches of the sky have received the same coverage.

%\subsection{Images \& Cross-matching}

@@

\begin{figure}[h!]
\centering
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs1.LOFARegs}.png}}
\caption{\label{fig.egs1.lofarim} Cutout of our LOFAR wide-field image centred on EGS-1.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs1.lofar}.png}}
\caption{\label{fig.egs1.sdss.lofoverlay} SDSS image of EGS-1, with \cref{fig.egs1.lofarim} as an overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs1.nvss}.png}}
\caption{\label{fig.egs1.sdss.nvssoverlay} Same image as \cref{fig.egs1.sdss.lofoverlay}, with NVSS overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs1.Wise3.4}.png}}
\caption{\label{fig.egs1.sdss.wiseoverlay} Same image as \cref{fig.egs1.sdss.lofoverlay}, with WISE overlay.}
\end{subfigure}
\caption{\label{fig.egs1} Images \& Overlays for EGS-1.}
\end{figure}

\pg
EGS-1 was selected from the wide-field LOFAR image because it showed interesting diffuse structure. It does not seem to match up to any optical source, nor does it match an infrared source. EGS-1 therefore seems to be best identified as a low-frequency counterpart to NVSS J143706+533225. Its shape seems to include two jets, one much brighter than the other, but both visible in \cref{fig.egs1.lofarim}. This is congruent with a radio galaxy with one Doppler-boosted jet. The absence of an optical counterpart could be due to dust extinction or to insufficient sensitivity.

\clearpage
\subsubsection{EGS-2}

\begin{figure}[h!]
\centering
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs2.LOFARegs}.png}}
\caption{\label{fig.egs2.lofarim} Cutout of our LOFAR wide-field image centred on EGS-2.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs2.lofar}.png}}
\caption{\label{fig.egs2.sdss.lofoverlay} SDSS image of EGS-2, with \cref{fig.egs2.lofarim} as an overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs2.nvss}.png}}
\caption{\label{fig.egs2.sdss.nvssoverlay} Same image as \cref{fig.egs2.sdss.lofoverlay}, with NVSS overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs2.Wise3.4}.png}}
\caption{\label{fig.egs2.sdss.wiseoverlay} Same image as \cref{fig.egs2.sdss.lofoverlay}, with WISE overlay.}
\end{subfigure}
\caption{\label{fig.egs2} Images \& Overlays for EGS-2.}
\end{figure}

\pg
EGS-2 was selected from the wide-field LOFAR image because of the interestingly turbulent morphology of its northern jet, and extended diffuse emission. 
A bright optical source lies very close to the mid-point of EGS-2. This source - 2MASX J14352846+5507519\footnote{\url{http://simbad.u-strasbg.fr/simbad/sim-id?Ident=\%40499314&Name=2MASX\%20J14352846\%2b5507519&submit=submit}} - is the brightest galaxy in a cluster. It lies at a redshift of 0.13985 \citepads{2013yCat.5139....0A}. Furthermore, there is an NVSS source which seems to nicely overlap the EGS-2: NVSS J143527+550756 \citepads{1998AJ....115.1693C}. These two sources do not seem to be associated in the literature, though 2MASX J14352846+5507519 is associated with VLSS 1435.4+5507 - a likely counterpart to the NVSS association. These are thus likely to all be the same object.

\pg
The northern edge of EGS-2's upper jet has interesting morphology: if EGS-2 is indeed associated to the brightest galaxy in the cluster, then this morphology is likely due to differing densities of the intercluster and intracluster media, which the jets will penetrate and interact with in different ways. As ever, because this source lies far from the observation pointing, the sensitivity and calibration are not optimal for analysis: further pointings on EGS-2 might be required to achieve the signal-to-noise required to extract useful information, especially using international LOFAR.

\subsubsection{EGS-3}

\begin{figure}[h!]
\centering
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs3.LOFARegs}.png}}
\caption{\label{fig.egs3.lofarim} Cutout of our LOFAR wide-field image centred on EGS-3.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs3.lofar}.png}}
\caption{\label{fig.egs3.sdss.lofoverlay} SDSS image of EGS-3, with \cref{fig.egs3.lofarim} as an overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs3.nvss}.png}}
\caption{\label{fig.egs3.sdss.nvssoverlay} Same image as \cref{fig.egs3.sdss.lofoverlay}, with NVSS overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs3.Wise3.4}.png}}
\caption{\label{fig.egs3.sdss.wiseoverlay} Same image as \cref{fig.egs3.sdss.lofoverlay}, with WISE overlay.}
\end{subfigure}
\caption{\label{fig.egs3} Images \& Overlays for EGS-3.}
\end{figure}

\pg
EGS-3 is one of the most impressive sources in the primary beam. It has very complex diffuse structure. It almost certainly consists of two jets, though the interactions of these jets with the neighbouring medium must be complex indeed. From \cref{fig.egs3.sdss.nvssoverlay}, one can see that the emission can be successfully matched to no less than two individual NVSS sources. In other words, if those NVSS sources are linked to the LOFAR source - and given their spatial distribution, they almost certainly are - then they are two components of a single, deeper structure. The infrared image tells us very little.

@@

\begin{figure}[h!]
\centering
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs4.LOFARegs}.png}}
\caption{\label{fig.egs4.lofarim} Cutout of our LOFAR wide-field image centred on EGS-4.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs4.lofar}.png}}
\caption{\label{fig.egs4.sdss.lofoverlay} SDSS image of EGS-4, with \cref{fig.egs4.lofarim} as an overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs4.nvss}.png}}
\caption{\label{fig.egs4.sdss.nvssoverlay} Same image as \cref{fig.egs4.sdss.lofoverlay}, with NVSS overlay.}
\end{subfigure}
\hfill
\begin{subfigure}{.48\textwidth}
\resizebox{\hsize}{!}{\includegraphics{images/{egs4.Wise3.4}.png}}
\caption{\label{fig.egs4.sdss.wiseoverlay} Same image as \cref{fig.egs4.sdss.lofoverlay}, with WISE overlay.}
\end{subfigure}
\caption{\label{fig.egs4} Images \& Overlays for EGS-4.}
\end{figure}

@@

\pg
The work done in this section had the primary goal of creating a decent model of the sky seen by LOFAR's primary beam. To do this, a wide-field, low-resolution image of the fully primary beam was made: this image makes no use of the LOFAR international stations due to technical limitations. Using PyBDSM \citepads{2015ascl.soft02007M}, this model was extracted from an image made after a few rounds of direction-dependent self-calibration.
